Specification of Advanced Types Translation

### 2014-03-08 DOCUMENTATION/AdvancedTypes.tex
Removed unneeded identifier and other small fix.

        \B{rhs} & : & \B{Expr} \\

@@

	\multicolumn{3}{|c|}{Label value: \B{LValue}} \\	

@@

        \B{inp} & : & seq \B{Expr} \\

@@

        \B{expr} & : & \B{Expr} \\

@@

        \multicolumn{3}{|c|}{Expression: \B{Expr}} \\

@@

\begin{tabular}[t]{rcl}
	\hline
        \multicolumn{3}{|c|}{Comparison: \B{Comp}} \\
        \hline
        \B{op} & : & \B{RelOp} \\
        \B{arg1}, \B{arg2} & : & \B{Exp} \\
        \hline
        \multicolumn{3}{|c|}{Data type: \B{Type}} \\
        \hline
        \multicolumn{3}{c}{\B{INT} \alt \B{BOOL}\alt}\\
	\multicolumn{3}{c}{\B{Struct} \alt \B{ArrayType}}\\
	\hline
  	\multicolumn{3}{|c|}{Data struct : \B{Struct}} \\
        \hline
	\B{id} & : &  \B{Name}\\
        \B{elements} & : & seq \B{Name} \\
	\B{type} & : & seq \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Access struct : \B{StructItem}} \\
        \hline
	\B{base} & : &  \B{Struct} \\
	\B{element} & : & \B{Name}\\
	\B{type} & : & \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Record struct : \B{Record}} \\
        \hline
	\B{elements} & : & seq \B{Name}\\
	\B{values} & : & seq \B{Expr}\\
	\hline
        \multicolumn{3}{|c|}{Array type : \B{ArrayType}} \\
        \hline 
	\B{dom} & : & seq \B{Set}\\
	\B{ran} & : & \B{Set}\\
        \hline
        \multicolumn{3}{|c|}{Array expression : \B{Array}} \\
        \hline
	\multicolumn{3}{c}{ seq \B{Map} \alt \B{Product} \alt \B{Name}  }\\
	\hline
  	\multicolumn{3}{|c|}{Ordered pair : \B{Map}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 
	\hline
  	\multicolumn{3}{|c|}{Cartesian product : \B{Product}} \\
        \hline
	\B{arg1} & : & seq \B{Expr}\\ 
	\B{arg2} & : & \B{Expr}\\ 	
	\hline
	\multicolumn{3}{|c|}{Access array : \B{ArrayItem}} \\
        \hline
	\B{base} & : &  \B{Array} \\
	\B{indexes} & : & seq \B{Expr}\\
	\B{type} & : & \B{ArrayType}\\
	\hline
	\multicolumn{3}{|c|}{Simple set : \B{Set}} \\
	\hline
	\multicolumn{3}{c}{\B{BOOL} \alt \B{NAT} \alt \B{NAT$_{1}$}\alt } \\
	\multicolumn{3}{c}{\B{INT} \alt \B{Interval}} \\
	\hline
	\multicolumn{3}{|c|}{B0 interval : \B{Interval}} \\
	\hline
	\B{start}, \B{end} & : & \B{Expr} \\
	\hline
	\multicolumn{3}{|c|}{Literal : \B{Lit}} \\
	\hline
	\multicolumn{3}{c}{ \B{Bool\_lit} \alt \B{Int\_lit}} \\ 
      \end{tabular}
    \end{tabular}
  }
  }
    \caption{Abstract syntax structure of a B implementation.  For each abstract
      syntax element, we give the name of the class (e.g. \B{CaseBr} for a
      branch in a case instruction, and the different attributes of the class),
      or a list of the possible sub-classes (e.g. \B{Inst} for the different
      kinds of instructions).
      Sequence and optional attributes are denoted ``seq'' and ``opt'',
      respectively.}
    \label{tab:node-attr}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{rcl}
      \nt{module} & ::= & \ListOf{\nt{item}} \\
      \nt{item} & ::= & \nt{const\_decl} \alt \nt{function\_decl} \alt \nt{type\_def}
      \alt \nt{const\_def} \alt \nt{var\_def} \alt \nt{function\_def} \\
      \nt{const\_decl} & ::= & \nt{name} \llvm{=} \llvm{external} \llvm{constant} \nt{type} \\
      \nt{type\_def} & ::= & \nt{name} \llvm{=} \llvm{type} \nt{type} \\
      \nt{type} & ::= & \llvm{void} \alt \nt{itype} \alt \llvm{\{} \ListOf{\nt{type}} \llvm{\}} \alt \nt{type}\llvm{*} \\
      \nt{const\_def} & ::= & \nt{name} \llvm{=} \llvm{constant} \nt{type} \nt{iliteral} \\
      \nt{var\_def} & ::= & \nt{name} \llvm{=} \llvm{common} \llvm{global} \nt{type} \llvm{zeroinitializer} \\
      \nt{function\_decl} & ::= & \llvm{declare} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{type}} \llvm{)}\\
      \nt{function\_def} & ::= & \llvm{define} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{param}} \llvm{)} \llvm{\{} \ListOf{\nt{block}} \llvm{\}} \\
      \nt{param} & ::= & \nt{type} \nt{name} \\
      \nt{block} & ::= & \nt{lbl} \llvm{:} \ListOf{\nt{inst}} \\
      \nt{inst} & ::=  & \nt{name} \llvm{=} \llvm{alloca} \nt{type} \\
      & \alt & \nt{name} \llvm{=} \lalt \llvm{add} \alt \llvm{sub} \alt \llvm{mul} \alt \llvm{sdiv} \alt \llvm{srem} \ralt \nt{itype} \nt{exp} \llvm{,} \nt{exp} \\
      & \alt & \nt{name} \llvm{=} \llvm{icmp} \lalt \llvm{eq} \alt \llvm{ne} \alt \llvm{sgt} \alt \llvm{sge} \alt \llvm{slt} \alt \llvm{sle} \ralt \llvm{i1} \nt{exp} \llvm{,} \nt{exp}\\
      & \alt & \nt{name} \llvm{=} \llvm{call} \nt{type} \llvm{(} \ListOf{\nt{arg}} \llvm{)} \\
      & \alt & \nt{name} \llvm{=} \llvm{getelementptr} \nt{type} \llvm{*} 
\nt{exp}\llvm{,} \nt{index}\llvm{,} \ListOf{\nt{index}} \\
      & \alt & \nt{name} \llvm{=} \llvm{load} \nt{type} \nt{exp} \\
      & \alt & \llvm{store} \nt{type} \nt{exp}, \nt{type} \llvm{*} \nt{exp} \\
      & \alt & \llvm{br} \llvm{i1} \nt{exp} \llvm{,} \llvm{label} \nt{lbl} \llvm{,} \llvm{label} \nt{lbl} \\
      & \alt & \llvm{br} \llvm{label} \nt{lbl} \\
%       & \alt & \llvm{switch} \nt{type} \nt{exp} \llvm{,} \llvm{branch} \nt{lbl} \llvm{[} \ListOf{\nt{branch}} \llvm{]} \\
      & \alt & \llvm{ret} \lalt \nt{type} \nt{exp} \alt \llvm{void} \ralt \\
      \nt{exp} & ::= & \nt{name} \alt \nt{iliteral} \alt \llvm{getelementptr} 
\llvm{(} \nt{type} \nt{exp} \llvm{,} \nt{index} \llvm{,} \ListOf{\nt{index}} \llvm{)} \\
      \nt{index} & ::= & \nt{itype} \nt{iliteral} \\
      \nt{branch} & ::= & \nt{iliteral} \nt{iliteral} \nt{lbl} \\
      \nt{arg} & ::= & \nt{type} \nt{exp}
    \end{tabular}
  \end{center}
  \caption{Grammar of the target LLVM subset: \nt{itype}, \nt{iliteral}, \nt{lbl}
    and \nt{name} correspond respectively to integer types, integer literals,
    labels and names. Choices are separated by \alt and optionally delimited by
    \lalt and \ralt.  The \ListOf{} superscript denotes a comma-separated list of
    elements of the annotated entity. }
  \label{fig:llvm-grammar}
\end{figure}

@@

\begin{align*}
\begin{split}
  \trad{a}{\B{Beq}} \isdef
  & \LET l \sep p \sep t'= \trad{a.lhs}{lvalue} \AND r \sep v \sep t = \trad{a.rhs}{\B{Expr}} \IN \\
  & \llvm{"\PH{l}} \\
  & \llvm{\PH{r}} \\
  & \llvm{store \PH{t} \PH{v}, \PH{t'} \PH{p} \nl"}
\end{split}
\end{align*}

@@

\begin{align*}
\begin{split}
  \trad{n}{\B{StructItem}} \isdef \\
   & \LET \ell = \lbl  \AND  \PH{t} = n.type \AND \PH{v} = n.element \IN\\
   & \llvm{"\PH{\ell} = getelementptr \%\PH{t}* @\PH{v}, i32 0, i32  \SBPos(\B{n})\nl"} \\
%   & \llvm{\PH{\ell2}  = load  \PH{\SBType(\B{n})}* \PH{\ell1} \nl"} \\
   & \sep  \ell  \sep \SBType(\B{n})\\
\end{split}
\end{align*}

>>>>

        \B{rhs} & : & \B{Exp} \\

@@

	\multicolumn{3}{|c|}{Left value: \B{LValue}} \\	

@@

        \B{inp} & : & seq \B{Exp} \\

@@

        \B{expr} & : & \B{Exp} \\

@@

        \multicolumn{3}{|c|}{Expression: \B{Exp}} \\

@@

\begin{tabular}[t]{rcl}
	\hline
        \multicolumn{3}{|c|}{Comparison: \B{Comp}} \\
        \hline
        \B{op} & : & \B{RelOp} \\
        \B{arg1}, \B{arg2} & : & \B{Exp} \\
        \hline
        \multicolumn{3}{|c|}{Data type: \B{Type}} \\
        \hline
        \multicolumn{3}{c}{\B{INT} \alt \B{BOOL}\alt}\\
	\multicolumn{3}{c}{\B{StructType} \alt \B{ArrayType}}\\
	\hline
  	\multicolumn{3}{|c|}{Data struct : \B{StructType}} \\
        \hline
        \B{elements} & : & seq \B{Name} \\
	\B{type} & : & seq \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Access struct : \B{StructItem}} \\
        \hline
	\B{base} & : &  \B{Name} \\
	\B{element} & : & seq \B{Name}\\
	\B{type} & : & \B{Type}\\
	\hline
	\multicolumn{3}{|c|}{Record struct : \B{Record}} \\
        \hline
	\B{elements} & : & seq \B{Name}\\
	\B{values} & : & seq \B{Exp}\\
	\hline
        \multicolumn{3}{|c|}{Array type : \B{ArrayType}} \\
        \hline 
	\B{dom} & : & seq \B{Set}\\
	\B{ran} & : & \B{Set}\\
        \hline
        \multicolumn{3}{|c|}{Array expression : \B{Array}} \\
        \hline
	\multicolumn{3}{c}{ seq \B{Map} \alt \B{Product} \alt \B{Name}  }\\
	\hline
  	\multicolumn{3}{|c|}{Ordered pair : \B{Map}} \\
        \hline
	\B{arg1} & : & seq \B{Exp}\\ 
	\B{arg2} & : & \B{Exp}\\ 
	\hline
  	\multicolumn{3}{|c|}{Cartesian product : \B{Product}} \\
        \hline
	\B{arg1} & : & seq \B{Exp}\\ 
	\B{arg2} & : & \B{Exp}\\ 	
	\hline
	\multicolumn{3}{|c|}{Access array : \B{ArrayItem}} \\
        \hline
	\B{base} & : &  \B{Array} \\
	\B{indexes} & : & seq \B{Exp}\\
	\B{type} & : & \B{ArrayType}\\
	\hline
	\multicolumn{3}{|c|}{Simple set : \B{Set}} \\
	\hline
	\multicolumn{3}{c}{\B{BOOL} \alt \B{NAT} \alt \B{NAT$_{1}$}\alt } \\
	\multicolumn{3}{c}{\B{INT} \alt \B{Interval}} \\
	\hline
	\multicolumn{3}{|c|}{B0 interval : \B{Interval}} \\
	\hline
	\B{start}, \B{end} & : & \B{Exp} \\
	\hline
	\multicolumn{3}{|c|}{Literal : \B{Lit}} \\
	\hline
	\multicolumn{3}{c}{ \B{Bool\_lit} \alt \B{Int\_lit}} \\ 
      \end{tabular}
    \end{tabular}
  }
  }
    \caption{Abstract syntax structure of a B implementation.  For each abstract
      syntax element, we give the name of the class (e.g. \B{CaseBr} for a
      branch in a case instruction, and the different attributes of the class),
      or a list of the possible sub-classes (e.g. \B{Inst} for the different
      kinds of instructions).
      Sequence and optional attributes are denoted ``seq'' and ``opt'',
      respectively.}
    \label{tab:node-attr}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{rcl}
      \nt{module} & ::= & \ListOf{\nt{item}} \\
      \nt{item} & ::= & \nt{const\_decl} \alt \nt{function\_decl} \alt \nt{type\_def}
      \alt \nt{const\_def} \alt \nt{var\_def} \alt \nt{function\_def} \\
      \nt{const\_decl} & ::= & \nt{name} \llvm{=} \llvm{external} \llvm{constant} \nt{type} \\
      \nt{type\_def} & ::= & \nt{name} \llvm{=} \llvm{type} \nt{type} \\
      \nt{type} & ::= & \llvm{void} \alt \nt{itype} \alt \llvm{\{} \ListOf{\nt{type}} \llvm{\}} \alt \nt{type}\llvm{*} \\
      \nt{const\_def} & ::= & \nt{name} \llvm{=} \llvm{constant} \nt{type} \nt{iliteral} \\
      \nt{var\_def} & ::= & \nt{name} \llvm{=} \llvm{common} \llvm{global} \nt{type} \llvm{zeroinitializer} \\
      \nt{function\_decl} & ::= & \llvm{declare} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{type}} \llvm{)}\\
      \nt{function\_def} & ::= & \llvm{define} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{param}} \llvm{)} \llvm{\{} \ListOf{\nt{block}} \llvm{\}} \\
      \nt{param} & ::= & \nt{type} \nt{name} \\
      \nt{block} & ::= & \nt{lbl} \llvm{:} \ListOf{\nt{inst}} \\
      \nt{inst} & ::=  & \nt{name} \llvm{=} \llvm{alloca} \nt{type} \\
      & \alt & \nt{name} \llvm{=} \lalt \llvm{add} \alt \llvm{sub} \alt \llvm{mul} \alt \llvm{sdiv} \alt \llvm{srem} \ralt \nt{itype} \nt{exp} \llvm{,} \nt{exp} \\
      & \alt & \nt{name} \llvm{=} \llvm{icmp} \lalt \llvm{eq} \alt \llvm{ne} \alt \llvm{sgt} \alt \llvm{sge} \alt \llvm{slt} \alt \llvm{sle} \ralt \llvm{i1} \nt{exp} \llvm{,} \nt{exp}\\
      & \alt & \nt{name} \llvm{=} \llvm{call} \nt{type} \llvm{(} \ListOf{\nt{arg}} \llvm{)} \\
      & \alt & \nt{name} \llvm{=} \llvm{getelementptr} \nt{type} \llvm{*} 
\nt{exp}\llvm{,} \nt{index}\llvm{,} \ListOf{\nt{index}} \\
      & \alt & \nt{name} \llvm{=} \llvm{load} \nt{type} \nt{exp} \\
      & \alt & \llvm{store} \nt{type} \nt{exp}, \nt{type} \llvm{*} \nt{exp} \\
      & \alt & \nt{name} \llvm{=} \llvm{bitcast} \nt{type} \nt{name}, \nt{type} \\
      & \alt & \llvm{br} \llvm{i1} \nt{exp} \llvm{,} \llvm{label} \nt{lbl} \llvm{,} \llvm{label} \nt{lbl} \\
      & \alt & \llvm{br} \llvm{label} \nt{lbl} \\
%       & \alt & \llvm{switch} \nt{type} \nt{exp} \llvm{,} \llvm{branch} \nt{lbl} \llvm{[} \ListOf{\nt{branch}} \llvm{]} \\
      & \alt & \llvm{ret} \lalt \nt{type} \nt{exp} \alt \llvm{void} \ralt \\
      \nt{exp} & ::= & \nt{name} \alt \nt{iliteral} \alt \llvm{getelementptr} 
\llvm{(} \nt{type} \nt{exp} \llvm{,} \nt{index} \llvm{,} \ListOf{\nt{index}} \llvm{)} \\
      \nt{index} & ::= & \nt{itype} \nt{iliteral} \\
      \nt{branch} & ::= & \nt{iliteral} \nt{iliteral} \nt{lbl} \\
      \nt{arg} & ::= & \nt{type} \nt{exp}
    \end{tabular}
  \end{center}
  \caption{Grammar of the target LLVM subset: \nt{itype}, \nt{iliteral}, \nt{lbl}
    and \nt{name} correspond respectively to integer types, integer literals,
    labels and names. Choices are separated by \alt and optionally delimited by
    \lalt and \ralt.  The \ListOf{} superscript denotes a comma-separated list of
    elements of the annotated entity. }
  \label{fig:llvm-grammar}
\end{figure}

@@

\begin{align*}
\begin{split}
  \trad{a}{\B{Beq}} \isdef
  & \LET l \sep p \sep t'= \trad{a.lhs}{lvalue} \AND r \sep v \sep t = \trad{a.rhs}{\B{Exp}} \IN \\
  & \llvm{"\PH{l}} \\
  & \llvm{\PH{r}} \\
  & \llvm{store \PH{t} \PH{v}, \PH{t'} \PH{p} \nl"}
\end{split}
\end{align*}

@@

\begin{align*}
\begin{split}
  \trad{n}{\B{StructItem}} \isdef \\
   & \IF \B{n.type} \mbox{ is \B{BOOL or NAT or INT} } \THEN \\
   &\quad  \LET \ell = \lbl  \AND  \PH{t} = n.type \AND \PH{v} = n.element \IN\\
   &\quad  \llvm{"\PH{\ell} = getelementptr \%\PH{t}* @\PH{v}, i32 0, i32  \SBPos(\B{n})\nl"} \\
   &\quad  \sep  \ell  \sep \SBType(\B{n})\\
\end{split}
\end{align*}
